    \documentclass[10pt,twocolumn]{article}
    \usepackage{fancyhdr, amsmath, amsthm, amssymb, mathtools, lastpage, enumerate, graphicx, setspace, wasysym, upgreek, listings}
    \usepackage[margin=0.5in, top=0.8in,bottom=0.8in]{geometry}
    \usepackage{microtype}
    \usepackage[bookmarks, pdfauthor={Brian Kubisiak, Yubo Su}]{hyperref}
    \usepackage[parfill]{parskip}
    \newcommand{\scinot}[2]{#1\times10^{#2}}
    \newcommand{\bra}[1]{\left<#1\right|}
    \newcommand{\ket}[1]{\left|#1\right>}
    \newcommand{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand{\norm}[1]{\left|\left|#1\right|\right|}
    \newcommand{\abs}[1]{\left|#1\right|}
    \newcommand{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand{\tensor}[1]{\overleftrightarrow{#1}}
    \let\Re\undefined%
    \let\Im\undefined%
    \newcommand{\ang}[0]{\text{\AA}}
    \newcommand{\mum}[0]{\upmu\mathrm{m}}
    \DeclareMathOperator{\Re}{Re}
    \DeclareMathOperator{\Im}{Im}
    \DeclareMathOperator{\Log}{Log}
    \DeclareMathOperator{\Arg}{Arg}
    \DeclareMathOperator{\Tr}{Tr}
    \DeclareMathOperator{\E}{E}
    \DeclareMathOperator{\Var}{Var}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator{\sgn}{sgn}
    \newcommand{\expvalue}[1]{\left<#1\right>}
    \usepackage[labelfont=bf, font=scriptsize]{caption}\usepackage{tikz}
    \usepackage[font=scriptsize]{subcaption}
    \everymath{\displaystyle}
    \lstset{basicstyle=\ttfamily\footnotesize,frame=single,numbers=left}

\tikzstyle{circ} = [draw, circle, fill=white, node distance=3cm, minimum height=2em]

\hypersetup{pdftitle={CUDA-MHD}}

\begin{document}

\pagestyle{fancy}
\rhead{CUDA-MHD --- Brian Kubisiak, Yubo Su}
\cfoot{\thepage/\pageref{LastPage}}

\title{CUDA-MHD\@: A GPU-accelerated Magnetohydrodynamic Simulation Scheme}
\author{Brian Kubisiak and Yubo Su}
\date{May 12, 2016}
\maketitle

\section{Summary}

Magnetohydrodynamics (MHD) is the study of fluid flow subject to changing
magnetic fields. Its diverse range of applications range from laboratory
experiments to solar phenomena to supernovae, the spectacular deaths of stars
hundreds of times the size of our Sun. MHD is described by a set of highly
complex nonlinear partial differential equations that require numerical
simulations of high accuracy to capture all interesting phenomena, a
computationally expensive task. Nevertheless, the numerical solution is highly
parallelizable and is well-suited to acceleration via GPU computation.

\section{Background}

The ideal MHD equations with the assumption of magnetic permeability $\mu = 1$
consist of the following five conservation laws:
\begin{align}
    \vec{\nabla} \cdot \vec{B} &= 0 \\
    \pd{\rho}{t} + \vec{\nabla} \cdot \left( \rho \vec{v} \right) &= 0 \\
    \pd{(\rho \vec{v})}{t} + \vec{\nabla} \cdot \left( \rho \vec{v}\vec{v} -
    \vec{B} \vec{B} \right) + \vec{\nabla}P^* &= 0 \\
    \pd{\vec{B}}{t} - \vec{\nabla} \times \left( \vec{v} \times \vec{B} \right)
    &= 0 \\
    \pd{E}{t} + \vec{\nabla} \cdot \left( (E + P^*)\vec{v} - \vec{B}\left(
    \vec{B} \cdot \vec{v} \right) \right) &= 0
\end{align}
where $\rho$ is the mass density, $\vec{v}$ is the velocity field (and $\rho
\vec{v}$ carries interpretation momentum density), $\vec{B}$ is the magnetic
field, $E$ the total energy density and $P^* = P + {B^2}/{2}$ the total
pressure where $P = (\gamma - 1)\left( E - \rho {v^2} / {2} - {B^2} / {2}
\right)$ is the gas pressure that satisfies the equation of state. The objective
is then to propagate all of these variables forward in time from some initial
state $\rho_0, {(\rho \vec{v})}_0, \vec{B}_0, \vec{E}_0, E_0, P^*_0$ while
obeying the conservation laws.

To solve these equations, we refer to the watershed paper \emph{Efficient
magnetohydrodynamic simulations on graphics processing units with CUDA} (Wong et
al. 2011) in which the ideal MHD equation were first implemented on the CUDA
architecture. While CUDA has evolved greatly since then, the general techniques
of the paper remain relevant. In the intended numerical scheme, the magnetic
field is first held fixed while the fiuid variables are updated. An analogous
procedure holding the fluid variables fixed and freeing the field is then
performed to complete one time step. The detailed procedure whereby this
evolution is performed to second-order accuracy can be found in the Wong et al.\
paper and will be the strategy we pursue.

\section{Anticipated Challenges}

The code will be immensely challenging simply due to the complexity of the full
problem. Moreover, the highly nonlinear nature of the dynamics that give rise to
the actually interesting MHD phenomena mandates numerical schemes of greater
complexity and lower error than most simpler methods (e.g. Runge-Kutta is a
lower bound on the required accuracy). However, the fact that it is a problem
that has great interest in the scientific community means that there is ample
literature on the preferred way to surmount these difficulties, at least
computational obstacles.

There are also implementational difficulties that will arise when trying to use
CUDA to speed up the computation. Since numerical MHD is linear in the number of
timesteps, no speedup can be incurred unless all data are stored in GPU memory
(otherwise, the simple cost of copying data will significantly decrease the
performance gains of a GPU implementation). The logistics of this storage will
be non-trivial to navigate. Additionally, the large number of variables and data
that will be stored at each timestep present a considerable challenge to
organize in a GPU-friendly way, and since the accuracy of a simulation scales
with its resolution, a space-friendly algorithm will also need to be found (to
maximize the resolution). The compromise between space-friendly resolution and
time-friendly optimization will certainly be a hurdle during the implementation
as well, one that will differ for our implementation on personal computers from
those in literature run on supercomputers.

\section{Goals}

We will deliver a full codebase for numerical MHD simulation alongside a sample
simulation in which a supernova of a star with realistic initial conditions is
simulated. The code will be capable of modelling any phenomena with appropriate
choice of initial conditions.

\section{Timeline}

\begin{description}
    \item[Week 1 (5/13--5/20)] --- We will complete literature reading and
        implement a CPU version of the desired MHD code. This implementation
        will be maximally modularized such that any GPU-accelerated functions
        can simply be substituted in ``drag and drop'' fashion.

    \item[Week 2 (5/20--5/27)] --- We will design the memory layout for the
        various variables in a GPU-friendly way. We will also begin considering
        how functions can be parallelized, a necessary part of finding an
        efficient memory layout scheme.

    \item[Week 3 (5/27--6/3)] --- We will implement the functions and run the
        simulation and obtain \textbf{FREAKING AWESOME RESULTS}.
\end{description}

\end{document}

